The Skew-Normal Approximation\\of the Binomial Distribution

### 2011-02-23 project/presentation/method-of-moments.tex
Tweaks.

%% The Central Moments of the Binomial
\frame{ \frametitle{Method of Moments: Central Moments of the Binomial}
  Let's start with the binomial ...
}
\frame{ \frametitle{Method of Moments: Central Moments of the Binomial}
  Let $B \sim Bin(n,p$.
  \pause

@@

  Why?
  \pause
  Recall Property 3 ...
}
\frame{ \frametitle{Method of Moments: Deriving an Approximation}
  \begin{center}
    \includegraphics[width=\textwidth]{../images/binomial-sn-n25p005.png}
  \end{center}
  When $p < 0.5$:
  \pause
  \begin{itemize}[<+->]
    \item The binomial skews right (weight shifts left) and approaches $+|N(0,1)|$ $\lra$ $\lambda$ is positive
    \item $(1-2p)$ is positive
  \end{itemize}
}
\frame{ \frametitle{Method of Moments: Deriving an Approximation}
  \begin{center}
    \includegraphics[width=\textwidth]{../images/binomial-sn-n25p095.png}
  \end{center}
  When $p > 0.5$:
  \pause
  \begin{itemize}[<+->]
    \item The binomial skews left (weight shifts right) and approaches $-|N(0,1)|$ $\lra$ $\lambda$ is negative
    \item $(1-2p)$ is negative
  \end{itemize}
}
\frame{ \frametitle{Method of Moments: Deriving an Approximation}
  Once you have $\lambda$, solve for $\sigma$ and then $\mu$.
  \pause
  \begin{equation*}
    \uncover<+->{np(1-p) = \sigma^2 \left( 1 - \frac{2}{\pi} \cdot \frac{\lambda^2}{1 + \lambda^2} \right) \quad\Rightarrow\quad
                 \sigma = \sqrt{\frac{np(1-p)}{1 - \frac{2}{\pi} \cdot \frac{\lambda^2}{1 + \lambda^2}}}}
  \end{equation*}
  \begin{equation*}
    \uncover<+->{np = \mu + \sigma \cdot \sqrt{\frac{2}{\pi}} \cdot \frac{\lambda}{\sqrt{1 + \lambda^2}} \quad\Rightarrow\quad
                 \mu = np - \sigma \cdot \sqrt{\frac{2}{\pi}} \cdot \frac{\lambda}{\sqrt{1 + \lambda^2}}}
  \end{equation*}
}
\frame{ \frametitle{Method of Moments: Deriving an Approximation}
  When $p = 0.5$, the binomial is symmetric, so $\lambda=0$.
  \pause

@@

%% Restrictions
\frame{ \frametitle{Method of Moments: Restrictions}
  Unfortunately, though better than the normal approximation, our skew-normal method isn't universal.
}
\frame{ \frametitle{Method of Moments: Restrictions}
  To be able to solve for $\lambda$, we must restrict $n$ and $p$ by the following equation:
  \pause
  \begin{equation}
    np(1-p) \geq \frac{1}{(1-2p)^2} \label{eq:solving-the-restriction}
  \end{equation}
  \pause
  From \eqref{eq:solving-the-restriction}, we can answer two questions:
}
\frame{ \frametitle{Method of Moments: Restrictions}
  \vspace{5mm}
  One: Given $p$, what is the least $n$ necessary?
  \pause
  \begin{equation*}
    n \geq \frac{(1-2p)^2}{p(1-p)}
  \end{equation*}
  \pause
  \begin{center}
    \includegraphics[width=0.7\textwidth]{../images/restriction-least-n.png}
  \end{center}
}
\frame{ \frametitle{Method of Moments: Restrictions}
  \vspace{5mm}
  Two: Given $n$, what is the range of possible $p$'s?
  \pause
  \begin{equation*}
    \frac12 - \frac12 \sqrt{\frac{n}{n+4}} \; \leq \; p \; \leq \; \frac12 + \frac12 \sqrt{\frac{n}{n+4}}
  \end{equation*}
  \pause
  \begin{center}
    \includegraphics[width=0.7\textwidth]{../images/restriction-p-range.png}
  \end{center}
}

>>>>

%% The Central Moments of the Binomial
\frame{ \frametitle{Method of Moments: Central Moments of the Binomial}
  Let's start with the binomial ...
}
\frame{ \frametitle{Method of Moments: Central Moments of the Binomial}
  Let $B \sim Bin(n,p)$.
  \pause

@@

  Why?
  \pause
  Recall Property 3 ...
}
\frame{ \frametitle{Method of Moments: Deriving an Approximation}
  \begin{center}
    \includegraphics[width=\textwidth]{../images/binomial-sn-n25p005.png}
  \end{center}
  When $p < 0.5$:
  \pause
  \begin{itemize}[<+->]
    \item The binomial skews right (weight shifts left) and approaches $+|N(0,1)|$ $\lra$ $\lambda$ is positive
    \item $(1-2p)$ is positive
  \end{itemize}
}
\frame{ \frametitle{Method of Moments: Deriving an Approximation}
  \begin{center}
    \includegraphics[width=\textwidth]{../images/binomial-sn-n25p095.png}
  \end{center}
  When $p > 0.5$:
  \pause
  \begin{itemize}[<+->]
    \item The binomial skews left (weight shifts right) and approaches $-|N(0,1)|$ $\lra$ $\lambda$ is negative
    \item $(1-2p)$ is negative
  \end{itemize}
}
\frame{ \frametitle{Method of Moments: Deriving an Approximation}
  Once you have $\lambda$, solve for $\sigma$ and then $\mu$.
  \pause
  \begin{equation*}
    \uncover<+->{np(1-p) = \sigma^2 \left( 1 - \frac{2}{\pi} \cdot \frac{\lambda^2}{1 + \lambda^2} \right) \quad\Rightarrow\quad
                 \sigma = \sqrt{\frac{np(1-p)}{1 - \frac{2}{\pi} \cdot \frac{\lambda^2}{1 + \lambda^2}}}}
  \end{equation*}
  \begin{equation*}
    \uncover<+->{np = \mu + \sigma \cdot \sqrt{\frac{2}{\pi}} \cdot \frac{\lambda}{\sqrt{1 + \lambda^2}} \quad\Rightarrow\quad
                 \mu = np - \sigma \cdot \sqrt{\frac{2}{\pi}} \cdot \frac{\lambda}{\sqrt{1 + \lambda^2}}}
  \end{equation*}
}
\frame{ \frametitle{Method of Moments: Deriving an Approximation}
  When $p = 0.5$, the right hand side of \eqref{eq:solving-for-lambda}
  \begin{equation*}
    \frac{\left( 1 - \frac{2}{\pi} \cdot \frac{\lambda^2}{1+\lambda^2} \right)^3}{\frac{2}{\pi} \left( \frac{\lambda^2}{1+\lambda^2} \right)^3 \left( \frac{4}{\pi} - 1 \right)^2}
    = \frac{np(1-p)}{(1-2p)^2}
  \end{equation*}
  is undefined.
  \pause

  Uh-oh.
}
\frame{ \frametitle{Method of Moments: Deriving an Approximation}
  Fortunately, our intuition rescues us:
  \pause

  When $p=0.5$, the binomial is symmetric, so $\lambda$ should equal 0.
  \pause

@@

%% Restrictions
\frame{ \frametitle{Method of Moments: Restrictions}
  Unfortunately, though better than the normal approximation, our skew-normal method isn't universal.
}
\frame{ \frametitle{Method of Moments: Restrictions}
  To be able to solve for $\lambda$, we must restrict $n$ and $p$ by the following equation:
  \pause
  \begin{equation}
    np(1-p) \geq (1-2p)^2 \label{eq:solving-the-restriction}
  \end{equation}
  \pause
  From \eqref{eq:solving-the-restriction}, we can answer two questions:
}
\frame{ \frametitle{Method of Moments: Restrictions}
  \vspace{5mm}
  One: Given $p$, what is the least $n$ necessary?
  \pause
  \begin{equation*}
    n \geq \frac{(1-2p)^2}{p(1-p)}
  \end{equation*}
  \pause
  \begin{center}
    \includegraphics[width=0.7\textwidth]{../images/restriction-least-n.png}
  \end{center}
}
\frame{ \frametitle{Method of Moments: Restrictions}
  \vspace{5mm}
  Two: Given $n$, what is the range of possible $p$'s?
  \pause
  \begin{equation*}
    \frac12 - \frac12 \sqrt{\frac{n}{n+4}} \; \leq \; p \; \leq \; \frac12 + \frac12 \sqrt{\frac{n}{n+4}}
  \end{equation*}
  \pause
  \begin{center}
    \includegraphics[width=0.7\textwidth]{../images/restriction-p-range.png}
  \end{center}
}
